### 2015-03-13 Paper/Secciones/reconstruccion.tex

\section{Reconstrucción}

\subsection{Subsection Heading Here}
Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.