### 2015-03-26 Paper2/implementacion.tex

\section{Implementación}\label{implementacion}

Acá por lo menos el diagrama de bloques general