### 2015-06-06 Paper_english/conclusiones.tex
Primera versión traducción de conlusiones

\section{Conclusions}

An optical motion capture system based on markers was obtained in full,
from capture video of a person in a lab environment with the right conditions, manages to obtain the 3D position of markers in the body of the person, representing the movement with an accuracy of centimeter order.\\
\hspace*{0.5cm} Furthermore, the implementation separates each process step in different modules capable of operating independently, allowing work with any acquisition system that generates not strictly optical images.\\
\hspace*{0.5cm} When testing with real sequences from three cameras in a laboratory outside the capture hypothesis, problems occur in the stages of segmentation and reconstruction. Performing background subtraction and modifications reconstruction barely acceptable even better results are achieved. Verifying the strong impact of a poor methodology of captured on the further processing.\\
\hspace*{0.5cm}This work does contribute to reproducibility and design methodology in the area through a structured full motion capture system.