### 2015-06-07 Paper_english/abstract.tex
Correciones Abstract
Correcciones minimas de sintaxis

\begin{abstract}
%\boldmath
This paper proposes an optical motion capture system based on markers that facilitates biomechanical motion analysis on people.
The system consists of separate blocks, giving an overall solution which allows to easily modify or replace components.
Guidelines are described for generating a virtual laboratory from which to obtain a database with sequences of synthetic videos and explore the characteristics required for a laboratory based on optical capture of markers to facilitate processing.
A set of metrics is presented to measure the performance of each block and the overall system.
Tests on the software implemented reflected that it has an accuracy of centimeter on sequences obtained in controlled environments. These results are good considering that the algorithms used in each block are of low complexity and can be optimized.
\keywords{
	Biomechanic, calibration, segmentation, reconstruction, tracking, virtual laboratory.
}
\end{abstract}