### 2015-06-07 Paper_english/conclusiones.tex
correccion conclusiones

\section{Conclusions}

It could be obtained a complete optical motion capture system based on markers,
that starting from video captures of a person in a lab environment with specific conditions, manages to obtain the 3D position of markers in the body of that person, representing the movement with an accuracy in the order of centimeter.\\
\hspace*{0.5cm} Furthermore, the implementation separates each process step in different modules capable of operating independently, allowing work with any acquisition system that generates non strictly optical images.\\
\hspace*{0.5cm} On the other hand, system fails in the stages of segmentation and reconstruction when testing with real sequences from three cameras in a laboratory outside the capture hypothesis. Better results are achieved performing background subtraction and making modifications in reconstruction algorithm. Verifying the strong impact of a poor methodology of captured on the further processing.\\
\hspace*{0.5cm}This work does contribute to reproducibility and design methodology in the area through a structured full motion capture system.